### 2023-04-26 Manuscript/Draft1.Rmd
manuscript update with a few more results from presentation

A simple visual analysis of all error plots for each species reveals that VAST estimates tend to provide abundance estimates that are above the true model value, while the stratified mean estimates are, on average, too low. This can be further examined by viewing estimate/ratio ratio plots, where we divide the yearly abundance estimates by the true model value. In doing so we see that VAST estimates tend to remain closer to the desired value of 1 compared to stratified mean estimates, which can range as low as 0 and exhibit large yearly changes.

@@

As noted in results second, adding covariate information to our Yellowtail Flounder VAST models always decreased the error in our abundance estimates, while covariate information sometimes lead to an increase in the absolute error between the abundance estimate and true model abundance for Cod and Haddock. More specifically, when a reduced number of strata are sampled adding covariate information leads to a decrease in performance for fall estimates of Haddock and Cod. The decline in performance can be explained by a failure to sample the full spectrum of temperature values where the species exists, which leads to an incomplete estimation of the covariate response. 
Generally help, but dependent on appropriate covariate response. Yellowtail cases always had appropriate response and thus adding covariates always improved estimates. VAST showed decreased performance with Cod and Haddock when covariates added in fall because covariate response was not fully developed. (show plot comparing good vs bad response)

increasing temperature scenarios resulted in a larger increase in average error for VAST estimates compared to stratified mean estimates. When including covarites the increase is less than not including them, but stratified mean saw a lower increase in the average error seen in constand an dincreasing temperature. Shouldnt model based be more stable with respect to increasing temperature?

@@

```{r PopPct, echo=FALSE,out.width="95%",fig.show='hold', fig.align='center', fig.cap="Percent of each species in each strata for during survey weeks in our spatial simulations. All constant temperature scenarios follow the patterns on the left while increasing temperature scenarios follow the patterns on the right.",out.width="47.5%"}
knitr::include_graphics(c("Images/PercentPlots_ConPopConTemp_YT.png","Images/PercentPlots_ConPopIncTemp_YT.png"))
knitr::include_graphics(c("Images/PercentPlots_DecPopConTemp_Cod.png","Images/PercentPlots_DecPopIncTemp_Cod.png"))
knitr::include_graphics(c("Images/PercentPlots_IncPopConTemp_Had.png","Images/PercentPlots_IncPopIncTemp_Had.png"))
```

>>>>

A simple visual analysis of all error plots for each species reveals that VAST estimates tend to provide abundance estimates that are above the true model value, while the stratified mean estimates are, on average, too low. This can be further examined by viewing estimate/model ratio plots, where we divide the yearly abundance estimates by the true model value and plot them on the log scale. In doing so we see that VAST estimates tend to remain closer to the desired value of 1 compared to stratified mean estimates, which can range as low as 0 and exhibit large yearly changes. A representative example of this trend is shown in Figure \@ref(fig:RatioTrend)

@@

As noted in results second, adding covariate information to our Yellowtail Flounder VAST models always decreased the absolute error in the resulting abundance estimates, while adding covariate information leads to a decrease in performance for fall estimates of Haddock and Cod when a reduced number of strata are sampled. The decline in performance can be explained by a failure to sample the full spectrum of temperature values where the species exists, leading to an incomplete estimation of the covariate response. Figure \@ref(fig:strata-plot) shows the strata excluded from sampling and Figure \@ref(fig:CovResponse) depicts the resulting covariate response for Yellowtail Flounder and Haddock for the same population scenario. Although strata were excluded for both species, the samples for Yellowtail Flounder ranged across the species preferred values and thus fully formed the covariate response, while the Haddock covariate response did not contain enough samples from the lower temperature range due to the specific strata that were excluded from sampling (ie, northern strata). As a result, when projecting into those strata as seen in the lower portion of Figure \@ref(fig:CovResponse), Yellowtail estimates with covariate information approximate the correct trend despite these strata being excluded from sampling, while the Haddock estimates that include strata provide that wrong trend at the wrong magnitude.

Generally help, but dependent on appropriate covariate response. Yellowtail cases always had appropriate response and thus adding covariates always improved estimates. VAST showed decreased performance with Cod and Haddock when covariates added in fall because covariate response was not fully developed. (show plot comparing good vs bad response)

increasing temperature scenarios resulted in a larger increase in average error for VAST estimates compared to stratified mean estimates. When including covariates the increase is less than not including them, but stratified mean saw a lower increase in the average error seen in constant an decreasing temperature. Shouldn't model based be more stable with respect to increasing temperature?

@@

```{r PopPct, echo=FALSE,fig.show='hold', fig.align='center', fig.cap="Percent of each species in each strata for during survey weeks in our spatial simulations. All constant temperature scenarios follow the patterns on the left while increasing temperature scenarios follow the patterns on the right.",out.width="47.5%"}
knitr::include_graphics(c("Images/PercentPlots_ConPopConTemp_YT.png","Images/PercentPlots_ConPopIncTemp_YT.png"))
knitr::include_graphics(c("Images/PercentPlots_DecPopConTemp_Cod.png","Images/PercentPlots_DecPopIncTemp_Cod.png"))
knitr::include_graphics(c("Images/PercentPlots_IncPopConTemp_Had.png","Images/PercentPlots_IncPopIncTemp_Had.png"))
```

```{r CovResponse, echo=FALSE,out.width="105%",fig.show='hold', fig.align='center', fig.cap="Temperature covariate response plots and resulting population estimate for Yellowtail Flounder on the left and Haddock on the right. In both cases the spatial simulations that were surveyed had an increasing population over time, increasing average temperature, and certain strata excluded from sampling, as shown in Figure 1**check**.",out.width="90%"}
# knitr::include_graphics(c("Images/covariate_good.png","Images/covariate_bad.png"))
# knitr::include_graphics(c("Images/covariate_est.png"))

knitr::include_graphics(c("Images/covariate_plot.png"))

```{r RatioTrend, echo=FALSE,out.width="105%",fig.show='hold', fig.align='center', fig.cap="Representative example of typical ratio trend as shown on a log scale.",out.width="90%"}
# knitr::include_graphics(c("Images/covariate_good.png","Images/covariate_bad.png"))
# knitr::include_graphics(c("Images/covariate_est.png"))

knitr::include_graphics(c("Images/ratio_trend.png"))

### 2023-04-26 Manuscript/Draft1.tex
manuscript update with a few more results from presentation

A simple visual analysis of all error plots for each species reveals that VAST estimates tend to provide abundance estimates that are above the true model value, while the stratified mean estimates are, on average, too low. This can be further examined by viewing estimate/ratio ratio plots, where we divide the yearly abundance estimates by the true model value. In doing so we see that VAST estimates tend to remain closer to the desired value of 1 compared to stratified mean estimates, which can range as low as 0 and exhibit large yearly changes.

@@

As noted in results second, adding covariate information to our Yellowtail Flounder VAST models always decreased the error in our abundance estimates, while covariate information sometimes lead to an increase in the absolute error between the abundance estimate and true model abundance for Cod and Haddock. More specifically, when a reduced number of strata are sampled adding covariate information leads to a decrease in performance for fall estimates of Haddock and Cod. The decline in performance can be explained by a failure to sample the full spectrum of temperature values where the species exists, which leads to an incomplete estimation of the covariate response.
Generally help, but dependent on appropriate covariate response. Yellowtail cases always had appropriate response and thus adding covariates always improved estimates. VAST showed decreased performance with Cod and Haddock when covariates added in fall because covariate response was not fully developed. (show plot comparing good vs bad response)

increasing temperature scenarios resulted in a larger increase in average error for VAST estimates compared to stratified mean estimates. When including covarites the increase is less than not including them, but stratified mean saw a lower increase in the average error seen in constand an dincreasing temperature. Shouldnt model based be more stable with respect to increasing temperature?

@@

>>>>

A simple visual analysis of all error plots for each species reveals that VAST estimates tend to provide abundance estimates that are above the true model value, while the stratified mean estimates are, on average, too low. This can be further examined by viewing estimate/model ratio plots, where we divide the yearly abundance estimates by the true model value and plot them on the log scale. In doing so we see that VAST estimates tend to remain closer to the desired value of 1 compared to stratified mean estimates, which can range as low as 0 and exhibit large yearly changes. A representative example of this trend is shown in Figure \ref{fig:RatioTrend}

@@

As noted in results second, adding covariate information to our Yellowtail Flounder VAST models always decreased the absolute error in the resulting abundance estimates, while adding covariate information leads to a decrease in performance for fall estimates of Haddock and Cod when a reduced number of strata are sampled. The decline in performance can be explained by a failure to sample the full spectrum of temperature values where the species exists, leading to an incomplete estimation of the covariate response. Figure \ref{fig:strata-plot} shows the strata excluded from sampling and Figure \ref{fig:CovResponse} depicts the resulting covariate response for Yellowtail Flounder and Haddock for the same population scenario. Although strata were excluded for both species, the samples for Yellowtail Flounder ranged across the species preferred values and thus fully formed the covariate response, while the Haddock covariate response did not contain enough samples from the lower temperature range due to the specific strata that were excluded from sampling (ie, northern strata). As a result, when projecting into those strata as seen in the lower portion of Figure \ref{fig:CovResponse}, Yellowtail estimates with covariate information approximate the correct trend despite these strata being excluded from sampling, while the Haddock estimates that include strata provide that wrong trend at the wrong magnitude.

Generally help, but dependent on appropriate covariate response. Yellowtail cases always had appropriate response and thus adding covariates always improved estimates. VAST showed decreased performance with Cod and Haddock when covariates added in fall because covariate response was not fully developed. (show plot comparing good vs bad response)

increasing temperature scenarios resulted in a larger increase in average error for VAST estimates compared to stratified mean estimates. When including covariates the increase is less than not including them, but stratified mean saw a lower increase in the average error seen in constant an decreasing temperature. Shouldn't model based be more stable with respect to increasing temperature?

@@

{\centering \includegraphics[width=0.9\linewidth]{Images/covariate_plot}

\caption{Temperature covariate response plots and resulting population estimate for Yellowtail Flounder on the left and Haddock on the right. In both cases the spatial simulations that were surveyed had an increasing population over time, increasing average temperature, and certain strata excluded from sampling, as shown in Figure 1**check**.}\label{fig:CovResponse}
\end{figure}

{\centering \includegraphics[width=0.9\linewidth]{Images/ratio_trend}

\caption{Representative example of typical ratio trend as shown on a log scale.}\label{fig:RatioTrend}
\end{figure}
